---
title: "word lists"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(ggthemes)
library(ggpubr)
```

# Exploring word lists

## Brysbaert & Biemiller 2017

```{r}
bb <-  read_xlsx("../data/word_lists/Brysbaert&Biemiller2017.xlsx")
```

> Brysbaert, M., & Biemiller, A. (2017). Test-based age-of-acquisition norms for 44 thousand English word meanings. Behavior research methods, 49(4), 1520-1523.

Variables:

 $ WORD        : from Dale and O’Rourke (1981)
 $ MEANING     : from Dale and O’Rourke (1981)
 $ AoAtestbased: updated test-based AoA measures from Dale and O’Rourke (1981) - values are in grades
 $ AoArating   : from Kuperman et al. (2012) - values are in years
 $ LWV         : original AoA measures from Dale and O’Rourke (1981) - values are in grades
 $ CDI         : age at which word is learned (produced according to CDI) - values are in month
 $ Morr        : age at which children could name pictures according to Morrison et al. (1997) - values are in month
 $ Floc        :suggested grade to teach the word according to the Flocabulary website - values are in grades
 
 ### Explore Morisson data
 
```{r}
mor <- bb %>%
  filter(!is.na(Morr))%>%
  filter(!duplicated(WORD))%>%
  mutate(aoa = round(Morr/12,2))

```

```{r}
mor %>%
  group_by(aoa)%>%
  summarise(n = n())%>%
  ggplot(aes(x = aoa, y = n))+
  geom_bar(stat = "identity", alpha = .5, col = "black", fill = "firebrick")+
  theme_minimal()


mor%>%
  mutate(WORD = factor(WORD),
         WORD = fct_reorder(WORD, -aoa))%>%
  ggplot(aes(y = WORD, x = aoa, label = WORD))+
  geom_text(alpha = .5, size = 2)+
  facet_wrap(~aoa, scales = "free", nrow = 1)+
  theme_few()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
  
```


#### Realtion to CDI

```{r}
mor %>%
  ggplot(aes(x = CDI, y = Morr))+
  geom_abline(intercept = 0, slope = 1, lty = 2, col = "grey")+
  geom_point(alpha = 0.2)+
  theme_minimal()+
  stat_cor(method = "pearson")+
  xlim(0,90)+
  ylim(0,90)
```
 
 